formula1

### 2020-02-20 inst/formula1.Rmd
fixes #27

Each row is a driver and each column (after the first) a venue.  We
see that Hamilton, the first row, came second in Australia, first in
China, second in Bahrain, fourth in Russia, and so on.  In the first
column, we see the result from Australia (AUS) in which Hamilton came
second, Vettel first, Bottas third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.  I now define a bespoke R function that
translates data of this type into a log-likelihood function using the
`hyper2` package.  The function is similar to, but slightly different
from, the analysis in `eurovision.R`.

```{r,defF1}
F1_likelihood <- function(wiki_table){

  noscore <- c("Ret", "WD", "DNS", "DSQ", "DNP", "NC")
  first_and_last <- c(1,ncol(wiki_table))

  racers <- wiki_table[,1]
  venues <- colnames(wiki_table)[-first_and_last]

  ## Now create a numeric matrix, fmat.  Two steps: first, count
  ## any no-score as zero:

  f <- function(x){
    if(any(x %in% noscore)){x[x%in%noscore] <- 0}
    return(x)
  }

  jj <- apply(wiki_table,2,f)

  ## Second, convert to numeric and strip out names; transpose of
  ## wiki_table (because we want each row to be a venue):

  fmat <- matrix(as.numeric(jj[,-first_and_last]),byrow=TRUE,ncol=nrow(wiki_table)) 
  colnames(fmat) <- racers
  rownames(fmat) <- venues

  F1 <- hyper2(d=ncol(fmat))

  ## Now cycle through the rows; each row is a venue [voter]
  for(i in seq_len(nrow(fmat))){
    d <- fmat[i,,drop=TRUE]
    while(any(d>0)){
      eligible <- which(d>=0)

      ## The first choice among eligible players has +1 power on the
      ## numerator:
      F1[which(d==1)] %<>% inc

      ## denominator of all eligible players; power -1
      F1[eligible] %<>% dec

      ## once you've come first in the field, you are ineligible to be first again:
      d[d==1] <- -1

      ## everyone moves down the list, so who *was* in second place
      ## becomes first place, who *was* third place becomes second,
      ## and so on:
      d[d>0] %<>% dec

    } # while() loop closes
  } # i loop closes

  ## syntatic sugar:
  pnames(F1) <- racers

  return(F1)
}
```

We can apply this to the dataset:

```{r,applydata}
L2017 <- F1_likelihood(f2017)
head(L2017,3)
```

@@

```{r,fouryearsdata}
F1_2014 <- "formula1_2014.txt" %>% read.table(header=TRUE) %>% F1_likelihood
F1_2015 <- "formula1_2015.txt" %>% read.table(header=TRUE) %>% F1_likelihood
F1_2016 <- "formula1_2016.txt" %>% read.table(header=TRUE) %>% F1_likelihood
F1_2017 <- "formula1_2017.txt" %>% read.table(header=TRUE) %>% F1_likelihood
a <- list(F1_2014, F1_2015, F1_2016, F1_2017)
alldrivers <- all_pnames(a)

>>>>

Each row is a driver and each column (after the first) a venue.  We
see that Hamilton, the first row, came second in Australia, first in
China, second in Bahrain, fourth in Russia, and so on.  In the first
column, we see the result from Australia (AUS) in which Hamilton came
second, Vettel first, Bottas third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.  I now use bespoke R function
\code{ranktable2supp()}, that translates data of this type into a
log-likelihood.  The function is similar to, but slightly different
from, the analysis in `eurovision.R`.

```{r,applydata}
L2017 <- ranktable2supp(f2017)
head(L2017,3)
```

@@

```{r,fouryearsdata}
F1_2014 <- "formula1_2014.txt" %>% read.table(header=TRUE) %>% ranktable2supp
F1_2015 <- "formula1_2015.txt" %>% read.table(header=TRUE) %>% ranktable2supp
F1_2016 <- "formula1_2016.txt" %>% read.table(header=TRUE) %>% ranktable2supp
F1_2017 <- "formula1_2017.txt" %>% read.table(header=TRUE) %>% ranktable2supp
a <- list(F1_2014, F1_2015, F1_2016, F1_2017)
alldrivers <- all_pnames(a)
